\speculake: Hiding Malware in Speculative Execution

### 2018-12-14 paper/architecture.tex
limits pass

In Table~\ref{table:processors} we analyze the number of necessary training
indirect jumps various processors require to consistently ($> 95\%$) have the
payload program enter the speculative world at the chosen speculative entry
point in the trigger program. We found that 28 indirect jumps was sufficient for
our trigger program on each of the test processors to reliably ensure the
speculative execution began at the correct speculative entry point.

@@

We find that when executing speculatively, the number of instructions completed
has a soft limit and a hard limit. The duration in cycles applies a soft limit,
as shown in Figure~\ref{fig:spec-capacity-np} with the \texttt{idiv (r32)},
\texttt{mul}, and \texttt{aesdec} instructions which see a steep drop in in hit
fraction, but continue to have a non-zero signal.  With all instructions a hard
limit is imposed by the number of CPU micro-operations ($\mu$-ops) that compose
the individual instruction. Figure~\ref{fig:spec-capacity-np} demonstrates that
the number of $\mu$-ops of the instructions is the major limiting factor that
define an upper boundary of approximately 150 instructions\footnote{While
\texttt{nop} is able to execute up to the full 220 ROB capacity, instructions
that do useful work (and/or use multiple $\mu$-ops) cannot reach this limit. On
top of this data dependency and execution unit availability add noise further
complicating the exact number of instructions that can be executed
speculatively.}

>>>>

In Table~\ref{table:processors} we analyze the number of necessary training
indirect jumps various processors require to consistently ($>~95\%$) have the
payload program enter the speculative world at the chosen speculative entry
point in the trigger program. We found that 28 indirect jumps was sufficient for
our trigger program on each of the test processors to reliably ensure the
speculative execution began at the correct speculative entry point.

@@

We find that when executing speculatively, the number of instructions completed
has a soft limit and a hard limit. The duration in cycles applies a soft limit,
as shown in Figure~\ref{fig:spec-capacity-np} with the \texttt{idiv (32-bit)},
\texttt{mul}, and \texttt{aesdec} instructions. As we attempt to execute more
instructions speculatively, we see a steep drop in in the fraction
of trials that are able to signal via the cache-side channel. However, this speculative
hit fraction does not drop to zero until the later hard limit,
imposed by the number of CPU micro-operations ($\mu$-ops) composing those instructions.
Figure~\ref{fig:spec-capacity-np} demonstrates that
the number of $\mu$-ops of the instructions is the major limiting factor that
define an upper bound of approximately 150 instructions\footnote{While
\texttt{nop} is able to execute up to the full 220 ROB capacity, instructions
that do useful work (and/or use multiple $\mu$-ops) cannot reach this limit.
In addition, data dependency and execution unit availability add further
complications to modelling the exact number of instructions that can be executed
speculatively.}.

### 2018-12-14 paper/figures.tex
correcting my mistake about cycles
limits pass

\newcommand{\FigSpecMeasureNP}{
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
    \caption{\textbf{Speculative limits}\,---\,%
    To find the number of instructions we can execute speculatively we placed a signal gadget
    after an increasing number of instructions and measured the fraction of times the 
    loaded value was subsequently in cache. We identify two limitation on the number of instructions
    accomplishable speculatively -- cache latency, and the CPU's reorder buffer~(ROB) size. We find that the
    hit fraction for instructions with a ratio of cycle latency to $\mu$~ops greater than 1 undergo a
    steep drop to \~0.1 hit fraction before falling definitively to 0. The drop in hit fraction 
    indicates that the duration of the test in cycles is beyond the fist step in cache latency as shown in 
    Figure~\ref{fig:cache-miss} (approx. 200 cycles). Whether or not the instruction sequence reaches the 
    cycle count limit we see only a sharp drop in hit fraction to 0 indicating the limiting factor 
    is ROB size. For example, the \texttt{idiv (r32)} instruction we identify 18 as the maximum number of 
    instruction limited by ROB (not including ~15 instructions for signal gadget), with a soft limit at 
    12 instructions enforced by the cache latency. We measure the ROB size by timing two cache 
    accesses with a variable number of instructions between them and note that this aligns with 
    the measured maximum \texttt{nop} instrustions.
    }
    \label{fig:spec-capacity-np}
\end{figure*}
}

>>>>

\newcommand{\FigSpecMeasureNP}{
\begin{figure*}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/Speculative_measurements_no-parity.png}
    \caption{\textbf{Speculative limits}\,---\,% To find the number of
    We placed a memory read after an
        increasing number of (speculatively executed) instructions and measured the
        fraction of times the
        loaded value was subsequently in cache. This tells us the upper bound of
        instructions we can reliably execute speculatively. We identify two
        limitations on the speculative lifetime:
        cache miss latency resolving the speculative branch, and the
        CPU's reorder buffer~(ROB) size. We observe that different instructions have
        varying speculative limits: for example, a 32-bit \texttt{idiv} can
        complete only 18 instructions, as each instruction inserts 10-$\mu$-ops
        into the ROB, while cheaper instructions that use fewer $\mu$-ops can execute
        more instructions.}

        % I think below is too much detail for the caption; it's likely to
        % confuse more than illuminate. We can leave this info in the text, but
        % 1) I don't know that the figure definitively shows what we're saying
        % here, and 2) it buries the more important (almost obvious) takeaway:
        % some instructions can execute longer speculatively, and modern CPUs
        % are complicated!

    %    We find that the hit fraction for
    %instructions with a ratio of cycle latency to $\mu$~ops greater than 1
    %undergo a steep drop to $\approx$ 0.1 hit fraction before falling
    %definitively to 0. The drop in hit fraction indicates that the duration of
    %the test in cycles is beyond the first step in cache latency as shown in
    %Figure~\ref{fig:cache-miss} (approx. 300 cycles). Whether or not the
    %instruction sequence reaches the cycle count limit we see only a sharp drop
    %in hit fraction to 0 indicating the limiting factor is ROB size. For
    %example, we identify 18 as the maximum number of \texttt{idiv (r32)}
    %instructions computable speculatively, limited by ROB (not including
    %$\approx$ 15 instructions for signal gadget), with a soft limit at 12
    %instructions enforced by the cache latency. We measure the ROB size by
    %timing two cache accesses with a variable number of instructions between
    %them and note that this aligns with the measured maximum \texttt{nop}
    %instrustions.}
    \label{fig:spec-capacity-np}
\end{figure*}
}
